\documentclass[twoside,11pt]{article}

% Any additional packages needed should be included after jmlr2e.
% Note that jmlr2e.sty includes epsfig, amssymb, natbib and graphicx,
% and defines many common macros, such as 'proof' and 'example'.
%
% It also sets the bibliographystyle to plainnat; for more information on
% natbib citation styles, see the natbib documentation, a copy of which
% is archived at http://www.jmlr.org/format/natbib.pdf

\usepackage{jmlr2e}
%\usepackage{parskip}

% Definitions of handy macros can go here
\newcommand{\dataset}{{\cal D}}
\newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}
% Heading arguments are {volume}{year}{pages}{submitted}{published}{author-full-names}

% Short headings should be running head and authors last names
\ShortHeadings{95-845: AAMLP Proposal}{Gangawar, Rost and Setia}
\firstpageno{1}

\begin{document}

\title{Heinz 95-845: Project Proposal}

\author{\name Mridul Gangawar \email mgangawar/mgangawar@andrew.cmu.edu \\
       \addr Heinz College of Information Systems and Public Policy\\
       Carnegie Mellon University\\
       Pittsburgh, PA, United States \\
       \AND
       \name Lauren Rost \email lrost/lrost@address.edu \\
       \addr Heinz College of Information Systems and Public Policy\\
       Carnegie Mellon University\\
       Pittsburgh, PA, United States \\
       \AND
       \name Nikita Setia \email nikitas/nikitas@address.edu \\
       \addr Heinz College of Information Systems and Public Policy\\
       Carnegie Mellon University\\
       Pittsburgh, PA, United States}
\maketitle


\section{Project Details}


\subsection{Objectives}
The objective of this project proposal is to generate a proposal for your course project. It should be concise and describe the following components:
\begin{itemize}
\item The premise of the analysis and a description of an analytic framework that motivates the use of machine learning for your task
\item Presentation of machine learning techniques appropriate for the task
\item Description of the data
\item Description of possible limitations of the study
\item Description of the likely analysis outcomes and their impact.
\end{itemize}

\subsection{Parameters}
The project will be conducted in groups of 2-3.

The project you propose should be different from an existing analysis, including publicly available analyses and analyses from other class projects of yours. It is permissible to perform an analysis in data that warrants a secondary analysis. My guideline here is that the analysis must be greater than 50\% new. To get approval for these studies, please describe the existing project and highlight the difference and contribution of this class's project. Provide any relevant documents (proposals, manuscripts, and/or citations). If the project has overlap with work from another course, you must also provide documented approval from the other faculty member/research collaborator(s). 

Your team is free to use programming language(s) of your choosing, however, we may only be able to support your endeavors in R.

\section{Proposal Details (10 points)} \label{details}
Please provide information for the following fields. Your proposal write-up should be no more than 2 pages.

\subsection{What is your proposed analysis? What are the likely outcomes?}


\subsection{Why is your proposed analysis important?}

\subsection{How will your analysis contribute to existing work? Provide references, \emph{e.g.}, see: \cite{cite1}.}

\subsection{Describe the data. Where applicable, please also define Y outcome(s), U treatment, V covariates, and W population.}


\subsection{What evaluation measures are appropriate for the analysis? Which measures will you use?}


\subsection{What study design, pre-processing, and machine learning methods do you intend to use? Justify that the analysis is of appropriate size for a course project.}

\subsection{What are possible limitations of the study?}

\subsection{Who will use your analytic pipeline? In one or two sentences, describe an example of its use.}

\bibliography{sample.bib}
%\appendix
%\section*{Appendix A.}
%Some more details about those methods, so we can actually reproduce them.

\end{document}
